Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2025-01-13 paper.qmd
Tried to add long edges within C-shaped cluster. Found: With UMAP, the model generate long edges within the clusters, while with tSNE between the clusters. The resaon that I think is that the characteric of tSNE to prserve global distance. To keep the clusters close. With UMAP, if the clusters are close the global structure already preserved. But to preserve the local distances, and didn't make the edges of C-shaped cluster too large, it make the long edges to keep the distances.

@@

free(distance_points) /
  wrap_plots(trimesh_two_curvy, trimesh_two_curvy_removed2, trimesh_two_curvy_removed1, two_curvy_proj_umap_model_delaunay, two_curvy_proj_umap_model_benchmark2, two_curvy_proj_umap_first_model1, ncol = 3) +
  plot_annotation(tag_levels = list(c("(a)", "(b)")))

>>>>

two_curvy_proj_umap_first_model1_temp <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#000000") + #31a354
  geom_point(
    size = 0.5,
    alpha = 0.2,
    color = "#636363") +
  geom_segment(
    data=axes,
    aes(x=x1, y=y1, xend=x2, yend=y2),
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2), 
    label=rownames(axes),
    colour="grey50",
    size = 4) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.1, 0.1)) + 
  ylim(c(-0.1, 0.1)) +
  interior_annotation("c2", sc_ltr_pos) +
  theme(
    legend.position = "none"
  )

@@

free(distance_points) /
  wrap_plots(trimesh_two_curvy, trimesh_two_curvy_removed2, trimesh_two_curvy_removed1, two_curvy_proj_umap_model_delaunay, two_curvy_proj_umap_model_benchmark2, two_curvy_proj_umap_first_model1_temp, ncol = 3) +
  plot_annotation(tag_levels = list(c("(a)", "(b)")))
